Towards atomistic resolution structure of phosphatidylcholine headgroup and glycerol backbone at different ambient conditions\footnote{Publication about results presented in the NMRlipids project}

### 2015-05-13 HGmodelMANUSCRIPT/coverLETTER.tex

To Whom It May Concern,

Please find attached to this message a manuscript that we would like to submit for publication 
in \textit{Journal of the american chemical society} (JACS): 
`Towards atomistic resolution structure of phosphatidylcholine glycerol backbone and choline headgroup at different ambient conditions', 
by A. Botan et al. In the manuscript we combine extensive amount of experimental Nuclear Magnetic Resonance (NMR)
and molecular dynamics simulation data to understand the atomistic resolution structure of phospholipid molecules and their assemblies.

We believe that the manuscript is interesting for the wide and diverse contemporary audience of
JACS since the used combination of experiments and simulations is shown to be successfull in
resolving the atomistic resolution structure of biomolecules in various biologically relevant conditions.
The approach is demonstrated for lipids, however, the extension to, e.g. membrane proteins is straightforward. 
Such an atomistic resolution of  biomolecular structure
significantly advances the understanding in fundamental chemistry, molecular biology, and also supports the development of applications, 
for example new biomaterials or drug delivery systems.

The work has been conducted by using a novel open collaboration concept, and all the scientific contributions
are done publicly through the nmrlipids.blogspot.fi. Thus all the scientific content related to this work
(including raw data, discussions, manuscript drafts, etc.)
has been publicly available during the whole course of the project through the blog. This material and discussion possibility
will be kept available also after the publication. This approach has significantly increased and will increase the scientific
quality and impact of the work. In addition, we have shared all the simulation raw data in such a format
that it can be reanalyzed easily also for other purposes. In addition to the scientic content,
we believe that the success of this approach in the field of chemistry is highly interesting 
for the JACS audience.

The usage of molecular dynamics simulations and open collaboration are adequate for the
problem at hand. Indeed, such simulations are currently the most straightforward method to 
construct atomistic resolution structures to interpret NMR experiments, but
enormous amount of simulation work was required for this work. The only practical
solution to collect this data is the open collaboration.
As result, we have produced original chemical insight for the atomistic resolution 
structures of phospholipids in various biologically relevant conditions.

We hope that you can consider this manuscript to be published as an article in the Journal of the american chemical society. \\

>>>>

Dear Editor,

Please find attached to this message a manuscript entitled `Towards atomistic resolution structure of phosphatidylcholine 
glycerol backbone and choline headgroup at different ambient conditions' by A. Botan et al., which we would like to submit 
for consideration of publication in \textit{Journal of the American Chemical Society} (JACS). In the manuscript a large amount of
existing experimental Nuclear Magnetic Resonance (NMR) data is reviewed and combined with an exceptionally comprehensive 
collection of generated molecular dynamics (MD) simulation data in order to understand the atomistic resolution structure 
of biologically abundant phospholipid molecules and their assemblies.

We believe that the manuscript is interesting to the wide and diverse contemporary audience of
JACS as it demonstrates how the combination of experimental NMR data with MD simulations is shown 
to be successful in resolving the atomistic resolution structure of biomolecules in various biologically 
relevant conditions. The approach is demonstrated for lipids, however, the extension to, e.g. membrane 
proteins is straightforward. Such an atomistic resolution of biomolecular structure significantly advances 
the understanding in fundamental chemistry, molecular biology, and also supports the development of applications, 
for example new biomaterials or drug delivery systems.

In addition to its main message, the manuscript also provides an extensive comparison on the quality of the head group 
structures of numerous widely used lipid models in varying biologically relevant conditions. The behavior
of the lipid head group region plays a key role in multiple processes, such as partitioning of drugs or 
nanoparticles into membranes and interactions of peripheral proteins with membranes, both of which are 
currently intensively studied. Therefore the manuscript provides crucial information for a large group
of computational scientists who are struggling to select the best possible lipid model for the problem at hand.

The work has been conducted by using a novel open collaboration concept, and all the scientific contributions
are made publicly through the blog-based project web page located at \texttt{nmrlipids.blogspot.fi}. 
Thus all the scientific content related to this work (including raw data, discussions, manuscript drafts, etc.)
has been publicly available during the whole course of the project through this blog. This material and the possibility to 
discuss it on the blog page will remain available also after the publication. This approach has significantly 
increased and will further increase the scientific quality and impact of the work. In addition, we have shared 
all the simulation raw data in a format that allows easy reanalysis and therefore reuse of the data for other purposes. 
In addition to the scientic content, we believe that the success of this fresh approach in the field of chemistry 
is highly interesting to the audience of JACS.

The usage of molecular dynamics simulations and open collaboration are adequate for the
problem at hand. Indeed, such simulations are currently the most straightforward method to 
construct atomistic resolution structures to interpret NMR experiments, but
enormous amount of simulation work was required for this work. The only practical
solution to collect this data was the open collaboration concept.
As result, we have produced original chemical insight for the atomistic resolution 
structures of phospholipids in various biologically relevant conditions.

We hope that you can consider this manuscript to be published as an article in the Journal of the American Chemical Society. \\
